\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Open Refine Learning Journal   - Bart Wojcik}
\author{Bart Wojcik}
\date{12 September 2019 | \textbf{Updated 7, 8 October 2019}}
\begin{document}

\maketitle

\section{OpenRefine for Social Science Data}
\subsection{Setup}
\begin{enumerate}
    \item \textbf{Action: }Downloaded, extracted and attempted to run OpenRefine 3.2.
    \subitem \textbf{Expectation: }Software executes. 
    \subitem \textbf{Result: }Software did not execute, OpenRefine reports dependence on JRE.
    \item \textbf{Action: }Downloaded and installed Java Runtime Environment, attempted to run OpenRefine again. 
    \subitem \textbf{Expectation: }Software executes.
    \subitem \textbf{Result: }Command prompt window opened and eventually a web interface launched in my default browser. Software executed successfully.
    \item \textbf{Action: }Lesson concluded.
\end{enumerate}
\subsection{Introduction}
\begin{enumerate}
    \item \textbf{Action: }Completed reading.
    \item \textbf{Action: }Lesson concluded.
\end{enumerate}
\subsection{Working with OpenRefine}
\begin{enumerate}
    \item \textbf{Action: }Attempted to import the SAFI\verb|_|openrefine.csv file.
    \subitem \textbf{Expectation: } File imports and opens correctly.
    \subitem \textbf{Result: }Successfully imported the file. Played a bit with the preview by changing the delimiter type to see what a bad import would look like. Reverted to default CSV to proceed with work.
    \item \textbf{Action: }Attempted to created the project.
    \subitem \textbf{Expectation: }Project creates.
    \subitem \textbf{Result: }Project created.
    \item \textbf{Action: }Used a text facet on the village column as instructed. 
    \subitem \textbf{Expectation: }Text facet produces results prescribed in the exercise.
    \subitem \textbf{Result: }Found the following issues when sorting: inconsistent vocabulary use between 'Chirdozo' and 'Chirodzo', 'Ruaca' and 'Ruca', 'Ruaca - Nhamuenda' and 'Ruaca-Nhamuenda'. '49' does not appear to be the right data for this field.
    \item \textbf{Action: }Completed exercise.
    \subitem \textbf{Expectation: }Expecting my findings to largely correspond to results expected in exercise solution.
    \subitem \textbf{Results: } Found: 19 unique results for 'interview\verb|_|date', the column appears to be formatted as text, transformed to date as instructed, had a look at the available facet and chose 'Timeline' in the way of testing, arrived at a timeline graph. Found that most interview data had been collected in 11/2016.
    \item \textbf{Action: }Tested clustering as instructed. 
    \subitem \textbf{Expectation: }I am expecting that using all of the clustering methods will result in clean data that requires no further manual intervention.
    \subitem \textbf{Result: }Found no more clustering after using 'metaphone3'
    \item \textbf{Action: }Manually narrowed down to four clusters as instructed.
    \subitem \textbf{Expectation: }Expecting results prescribed in the exercise solution.
    \subitem \textbf{Result: }As prescribed in the exercise.
    \item \textbf{Action: }Moving to 'Transforming data'. 
    \item \textbf{Action: }Removed all left square brackets as instructed. Removed single quotes using the method prescribed in the exercise. Removed right square brackets.
    \subitem \textbf{Expectation: }Expecting output to match the results prescribed in the exercise
    \subitem \textbf{Result: }Output matches expected results.
    \item \textbf{Action: }Split items owned using a custom text facet.
    \subitem \textbf{Expectation: }Expecting results as prescribed in the exercise.
    \subitem \textbf{Result: }Noticed several duplicate labels. 'Mobile phone' and 'radio' appear to be the two most owned items; 'car' and 'computer' the least owned. The duplicate labels need clean-up before this step, why is this missing from the worksheet?!? I thought the idea was to get the machine to do the job not have me look for repeated instances of a label. November was the month most reported for food lack.
    \item \textbf{Action: }Performed \newline \verb|"value.replace("[", "").replace("]", "").replace("'", "").replace(" ", "")"| on \verb|months_no_water|, \verb|liv_owned|, \verb|res_change|, \verb|no_food_mitigation| by recycling the command using the history tab.
    \subitem \textbf{Expectation: }Expecting output without left and right square brackets or spaces on these columns.
    \subitem \textbf{Result: }Results as expected. Trimmed leading and trailing white-spaces and ended with only four labels as predicted by the exercise.
    \item \textbf{Action: }Scoped the undo\verb|/|redo feature. Undoing and redoing works as you would from undoing and redoing.
    \item \textbf{Action: }Concluded the exercise.
\end{enumerate}
\subsection{Filtering and Sorting with OpenRefine}
\begin{enumerate}
    \item \textbf{Action: }Completed the filtering section following the instructions outlined in the exercise.
    \subitem \textbf{Expectation: }Expecting output to match that prescribed in the exercise.
    \subitem \textbf{Result: } The selected types of roof were all the types that contained the string 'mabat', to further restrict the selection the user should provider a more closely phrased match.
    \item \textbf{Action: }Scoped the \verb|include\exclude| feature following the instructions outlined in the exercise.
    \subitem \textbf{Expectation: }I expect that using this feature will allow me to include or exclude parts of the data under investigation as required.
    \subitem \textbf{Result: }The feature works as expected.
    \item \textbf{Action: }Attempting to sort the tables as outlined in the exercise.
    \subitem \textbf{Expectation: }I expect that sorting on a column will sort the data contained in that column.
    \subitem \textbf {Result: }Sorting does nothing in my software, the command is simple and I have not made a mistake. I had to refresh the page for it to work again, I suspect my session either expired or timed out due to inactivity. To answer the question (asked by the lesson) of whether there is anything wrong, the lesson makes a tall assumption about my understanding of GPS based altitude recording, if I had to, I would say zero means 'sea level' but the lesson solution says that it means the device could not determine a value. Completed the sorting exercise and quickly discovered by looking at GPS coordinates that village 49 should be Chirodzo.
    \item \textbf{Action: }Changed the village name as instructed.
    \item \textbf{Action: }Lesson concluded.
\end{enumerate}
\subsection{Examining Numbers in OpenRefine}
\begin{enumerate}
    \item  \textbf{Action: }Execute a common transform to numerical values on the years\verb|_|farm column.
    \subitem \textbf{Expectation: }The cell data transforms from text strign to numerical value that allows math to be performed.
    \subitem \textbf{Result: }Success.
    \item \textbf{Action: }Applied the same common transform to no\verb|_|members, yrs\verb|_|liv and buildings\verb|_|in\verb|_|compound columns as instructed.
    \subitem \textbf{Expectation: }As above.
    \subitem \textbf{Result: }As above
    \subitem \textbf{Answer to exercise question: }Only cells containing numerical data can be transformed in this way.
    \item \textbf{Action: }Amended a cell previously transformed to store numerical data by spelling out the number.
    \subitem \textbf{Expectation: }The edit window data type drop down indicated the data will be stored as a text string so I expect it to be stored as such.
    \subitem \textbf{Result: }The value is indeed once more stored as a text string.
    \item \textbf{Action: }Applied text facet to the affected column as instructed.
    \subitem \textbf{Expectation: }I expect the text facet to report the spelled out number as a non-numeric value.
    \subitem \textbf{Result: }The spelled out value is indeed report as such and can be included or excluded from viewing using the non-numeric checkbox in this facet.
    \item \textbf{Action: }Lesson concluded.
\end{enumerate}
    \subsection{Using scripts}
    \begin{enumerate}
        \item \textbf{Action: }Exported operations executed so far as a script as per exercise instructions.
        \subitem \textbf{Expectation: }I expect that the software will save the history of work as a set of commands that can be re-imported and re-run on the data set, thus replicating the results I have achieved so far.
        \subitem \textbf{Results: }OpenRefine exported the history of the performed data cleanup as a huma-readable JSON file.
        \item \textbf{Action: }Re-imported the JSON script back into OpenRefine on a new, unmodified project based on the same SAFI source file.
        \subitem \textbf{Expectation: }OpenRefine re-executes the commands stored in the script resulting in an instant clean up of the unmodified data.
        \subitem \textbf{Result: }Results as expected.
        \item \textbf{Action: }Lesson concluded.
    \end{enumerate}
    \subsection{Exporting and Saving Data from OpenRefine}
    \begin{enumerate}
        \item \textbf{Action: }Exported the project to a local file as instructed.
        \subitem \textbf{Expectation: }I expect that OpenRefine will, in fact, export my data along with the history of operations performed on it to a local file.
        \subitem \textbf{Result: }OpenRefine has indeed done so. I was expecting a single file (rather than an archive) so I was surprised to see a .tar.gz file. I extracted it with 7zip and found that it contains everything I outlined in my expectations for it.
    \end{enumerate}
    \subsection{Other Resource in OpenRefine}
    \begin{enumerate}
        \item \textbf{Action: }Read through contained material.
        \subitem \textbf{Expectation: }I expect to find a bare modicum to no applicability of OpenRefine and data cleaning principles to the kind of qualitative research I am looking at carrying out at this tine,
        \subitem \textbf{Result: }Concluded reading, results as expected.
        \item \textbf{Action: }Concluded OpenRefine material.
    \end{enumerate}
\end{document}