\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Week 2 - Learning Journal 1 - Bart Wojcik}
\author{Bart Wojcik}
\date{August 2019}

\begin{document}

\maketitle

\section{14/08/2019 - Learning Journal}
\subsection{Getting started with Overleaf and Ocrid:}
\medskip
\begin{itemize}
    \item I've attempted to sing into Overleaf using ORCID to avoid creating a new set of logins for everything that can otherwise be connected in one place.
    \item Although ORCID sign-in took me to a MQU branded login portal and allowed me to configure personal data sharing permissions, I still had finalise account creation on ORCID itself. 
    \item I thought I'd update my profile while I'm there (the ORCID profiles seem to be public to some degree and I'm curious if I can add the Resurge Digital web address to a publicly visible profile, we can always use backlinks at Resurge, but I'm currently stuck on https://orcid.org/my-orcid watching two spinning arrows that clearly aren't loading the content that I requested even though they would like me to believe that they are). 
    \item Rather than reloading the profile management page, I opted to go back to the root domain at https://orcid.org/ and click through to https://orcid.org/my-orcid, this resolved my problem). 
    \item If this is going to be my long term research profile, I may as well take the time to clone my LinkedIn details. 
\end{itemize}
\subsection{Notes:}
I was happy to see that the company backlink that I added to the employment on OCRID section displays publicly (I verified that in Private Browsing) and that it does not include a rel="nofollow" attribute (the page does not include a <meta name="robots" content="noindex"> tag either, nor does the domain server seem to have https://orcid.org/robots.txt file with any 'disallow' directives that could prevent Googlebot from crawling my public profile page). The link in the general URLs section has a rel="nofollow" attribute automatically appended but the link in the 'employment' section does not (it is hidden in an accordion but I remember that John Mueller insisted on Googlebot indexing accordion content contrary to the popular belief amongs SEOs). There is still the issue of whether this profile page is publicly discover-able but I can point Google in this direction at a later date. While I was at it, I've added a backlink to my Github public profile too but this one (disappointingly) also had a nofollow attribute appended.

\end{document}
